---
title: "Urban Farm Dashboard - User Guide"
author: "Urban Farm Team"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Urban Farm Dashboard

## Overview

The Urban Farm Dashboard is a comprehensive RShiny application designed to help urban farmers manage their farm operations efficiently. The application focuses on two critical areas:

1. **Chicken Management**: Track and monitor your chicken flock's health, inventory, and characteristics
2. **Hazard Management**: Identify, log, and track farm hazards to ensure a safe working environment

## Installation

### Prerequisites

Ensure you have R (version 3.5.0 or higher) and RStudio installed on your system.

### Required Packages

Install the required R packages by running:

```{r install-packages}
install.packages(c(
  "shiny",
  "shinydashboard", 
  "DT",
  "ggplot2",
  "dplyr",
  "googlesheets4",
  "gargle"
))
```

### Running the Application

To run the application locally:

1. Clone or download this repository
2. Set your working directory to the project root
3. Run the following command:

```{r run-app}
shiny::runApp()
```

Alternatively, open the `app.R` file in RStudio and click the "Run App" button.

### Google Sheets Configuration (Optional)

To enable persistent storage in Google Sheets, set these environment variables:

```
GSHEETS_SHEET_ID=<your-sheet-id>
GSHEETS_CHICKENS_TAB=Chickens
GSHEETS_HAZARDS_TAB=Hazards
GSHEETS_SERVICE_JSON=<service-account-json>
```

If not set, the app falls back to CSV files under `data/`.

## Application Structure

### File Organization

```
Urban_farm_dashboard/
├── app.R                    # Main application file
├── DESCRIPTION             # Package dependencies
├── README.Rmd              # This documentation file
├── R/
│   └── helpers.R           # Helper functions
├── modules/
│   ├── chickens_module.R   # Chicken management module
│   └── hazards_module.R    # Hazard management module
├── data/
│   ├── chickens.csv        # Chicken inventory data
│   └── hazards.csv         # Hazard log data
└── www/
    └── custom.css          # Custom styling
```

### Architecture

The application follows a modular design pattern:

- **app.R**: Main entry point containing UI and server logic
- **modules/**: Shiny modules for different features (chickens, hazards)
- **R/helpers.R**: Utility functions for data management and validation
- **data/**: CSV files for persistent storage
- **www/**: Static assets (CSS, images, etc.)

## Features

### Dashboard

The main dashboard provides an overview of your farm:

- **Total Chickens**: Current count of chickens in inventory
- **Active Hazards**: Number of unresolved hazards
- **Tasks Today**: Scheduled tasks (future enhancement)

### Chicken Management

#### Viewing Chicken Inventory

The Chicken Management page displays:

- Complete inventory table with sorting and search capabilities
- Health status distribution chart
- Breed distribution chart

#### Adding New Chickens

To add a new chicken:

1. Navigate to the "Chickens" tab
2. Click "Add New Chicken"
3. Fill in the required information:
   - Name
   - Breed
   - Age (in months)
   - Health Status
   - Last Health Check Date
4. Click "Add Chicken"

#### Data Fields

- **ID**: Unique identifier (auto-generated)
- **Name**: Chicken's name
- **Breed**: Breed type (e.g., Rhode Island Red, Plymouth Rock)
- **Age (months)**: Age in months
- **Health Status**: Current health condition
  - Healthy
  - Sick
  - Recovering
  - Under Observation
- **Last Check**: Date of last health examination

### Hazard Management

#### Viewing Hazards

The Hazard Management page includes:

- Filterable hazard log table
- Severity distribution visualization
- Status overview chart

#### Reporting New Hazards

To report a new hazard:

1. Navigate to the "Hazards" tab
2. Click "Report New Hazard"
3. Provide details:
   - Hazard Type
   - Description
   - Severity Level
   - Status
   - Date Reported
4. Click "Report Hazard"

#### Hazard Types

- Animal Health
- Facility Structure
- Equipment Failure
- Weather Related
- Predator Activity
- Disease Outbreak
- Other

#### Severity Levels

- **Low**: Minor issue, low priority
- **Medium**: Moderate issue, should be addressed soon
- **High**: Serious issue, requires prompt attention
- **Critical**: Emergency situation, immediate action required

#### Hazard Status

- **Active**: Currently unresolved
- **Under Review**: Being investigated or planned
- **Resolved**: Issue has been addressed

#### Filtering Hazards

Use the filter controls to:

- View hazards by status (All, Active, Resolved, Under Review)
- Filter by severity level (All, Low, Medium, High, Critical)

## Data Management

### Data Storage

Data is stored in CSV files in the `data/` directory:

- `chickens.csv`: Chicken inventory
- `hazards.csv`: Hazard log

### Data Backup

It's recommended to regularly backup your data files. You can:

1. Copy the `data/` folder to a safe location
2. Use version control (Git) to track changes
3. Export data from the application (future feature)

### Data Validation

The application includes validation for:

- Required fields (name, breed, type, description)
- Numeric values (age must be positive)
- Severity levels (must be valid option)

## Customization

### Adding New Features

The modular architecture makes it easy to add new features:

1. Create a new module file in `modules/`
2. Define UI function: `moduleNameUI()`
3. Define server function: `moduleNameServer()`
4. Add module to `app.R`

### Styling

Custom CSS can be edited in `www/custom.css`. The application uses:

- Bootstrap 3 (via shinydashboard)
- Custom color scheme with green theme
- Responsive design for mobile devices

## Troubleshooting

### Common Issues

**App won't start**
- Check that all required packages are installed
- Verify that the working directory is correct
- Check R console for error messages

**Data not loading**
- Ensure CSV files exist in the `data/` directory
- Check file permissions
- Verify CSV format (comma-separated, proper headers)

**Changes not saving**
- Check write permissions on the `data/` directory
- Ensure files are not open in another program
- Verify disk space availability

## Future Enhancements

Planned features include:

- [ ] Task scheduling and tracking
- [ ] Export functionality for reports
- [ ] Email notifications for critical hazards
- [ ] Mobile app version
- [ ] Multi-user support with authentication
- [ ] Integration with IoT sensors
- [ ] Automated health check reminders
- [ ] Feed and supply inventory management
- [ ] Weather data integration
- [ ] Financial tracking

## Contributing

To contribute to this project:

1. Fork the repository
2. Create a feature branch
3. Make your changes
4. Submit a pull request

## Support

For questions or issues:

- Open an issue on GitHub
- Contact the farm management team
- Check documentation in `README.Rmd`

## Appendix: Documentation Source and Publishing

Documentation is managed in two layers:

1. **Source documentation** (editable project docs): `documentation/`
2. **Published documentation** (GitHub-facing output): `docs/`

After editing docs in `documentation/`, republish to `docs/` by running:

```{r publish-docs, eval=FALSE}
source("scripts/publish_docs.R")
```

## License

This project is licensed under the MIT License - see the LICENSE file for details.

## Acknowledgments

Built with:

- [Shiny](https://shiny.rstudio.com/) - Web Application Framework for R
- [shinydashboard](https://rstudio.github.io/shinydashboard/) - Dashboard Framework
- [DT](https://rstudio.github.io/DT/) - Interactive Tables
- [ggplot2](https://ggplot2.tidyverse.org/) - Data Visualization
- [dplyr](https://dplyr.tidyverse.org/) - Data Manipulation

---

*Last updated: `r Sys.Date()`*
