---
title: "Quick Start Guide"
output: html_document
---

# Quick Start Guide

## Getting Started in 5 Minutes

### 1. Install Required Packages

```{r eval=FALSE}
# Install dependencies
install.packages(c(
  "shiny",
  "shinydashboard", 
  "DT",
  "ggplot2",
  "dplyr",
  "googlesheets4",
  "gargle"
))
```

### 2. Run the Application

```{r eval=FALSE}
# Option 1: From R console (in the project directory)
setwd("~/Urban_farm_dashboard")
shiny::runApp()

# Option 2: From command line
Rscript run_app.R

# Option 3: From RStudio
# Open app.R and click "Run App" button
```

The app will automatically open in your browser at `http://127.0.0.1:XXXX`

### 3. Navigate the Dashboard

The dashboard has four main tabs in the sidebar:

- **Dashboard**: View overview statistics (total chickens, active hazards)
- **Chickens**: Manage your flock inventory and health records
- **Hazards**: Track and report farm hazards
- **Farm Risk** (NEW): KVC Risk calculator, profiles, and crowdsourced assessments
- **About**: Learn more about the application

## Key Features at a Glance

### Managing Chickens

1. Click "Chickens" in the sidebar
2. Click "Add New Chicken" button
3. Fill in chicken details (name, breed, age, health status)
4. Confirm and view in inventory table
5. Data is automatically saved to `data/chickens.csv`

### Reporting Hazards

1. Click "Hazards" in the sidebar
2. Click "Report New Hazard" button
3. Describe the hazard, set severity, and status
4. Submit to add to hazard log
5. Use filters to view hazards by status or severity

### Farm Risk Assessment (NEW)

#### Tab 1: Job Risk Calculator
1. Select a farm zone (e.g., "Sheep Yards", "Chicken Coop")
2. Check off all tools needed for the job
3. View:
   - Area Max Hazard Score
   - Highest Tool Risk Class
   - Operational Status (STANDARD / HIGH / CRITICAL)
   - Detailed hazard table with controls

#### Tab 2: Compare Risk Profiles
1. Select a tool from the dropdown
2. View comparison chart: Teacher Baseline vs AI Engine vs Historical vs Live Class
3. Check statistical variance table

#### Tab 3: Log Assessment
1. Select a tool being assessed
2. Move the risk slider (1–5)
3. Enter the spatial coordinate/landmark
4. Write your justification
5. Click "Submit To Master Registry"
6. View submission instantly in the live feed table

## Data & Offline Mode

**Data is automatically saved in two places:**
- Local CSV files in `data/` folder (always)
- Google Sheets (if configured with environment variables)

**If Google Sheets is unavailable**, the app automatically falls back to local CSV files — no data is lost.

### Setting Up Google Sheets (Optional)

See [google_sheets_setup.Rmd](google_sheets_setup.Rmd) for detailed instructions.

## Tips

- Use the **Refresh** button to reload data from storage
- Filter hazards by **status** and **severity** for better insights
- Check **visualizations** in each module for patterns
- Enable **offline mode** by not setting Google Sheets environment variables
- Data persists even if you restart the app

## Common Tasks

### Export Data

Data is stored as CSV files in the `data/` folder:
- `chickens.csv` – Chicken inventory
- `hazards.csv` – Hazard log
- `area_hazards.csv` – Hazard reference (read-only)
- `tool_registry.csv` – Tool risk registry (read-only)
- `risk_submissions.csv` – Student assessments

You can open these files in Excel or any text editor.

### Backup Data

```bash
# Create a backup folder
mkdir -p data/backups

# Backup important files
cp data/chickens.csv data/backups/chickens_$(date +%Y%m%d).csv
cp data/hazards.csv data/backups/hazards_$(date +%Y%m%d).csv
cp data/risk_submissions.csv data/backups/risk_submissions_$(date +%Y%m%d).csv
```

### Reset Data

To start fresh, delete the CSV files:

```bash
rm data/chickens.csv
rm data/hazards.csv
```

The app will create new empty files on next run.

## Troubleshooting

### App won't start

**Error: "package not found"**
- Install missing packages: `install.packages("package_name")`
- Run `Rscript run_app.R` to auto-install

**Error: "Port already in use"**
- Close other Shiny apps or wait 10 seconds, then try again
- Or specify a different port: `shiny::runApp(port = 3839)`

### Data not saving

- Check `data/` folder exists
- Verify file permissions allow writing
- Check the console for error messages

### Google Sheets not connecting

- Verify environment variables are set correctly
- Check service account has edit access to sheet
- App will automatically fall back to local CSV files

## Next Steps

1. **Read the full README**: See [../README.md](../README.md) for comprehensive documentation
2. **Check data schema**: See [data_schema.md](../resources/data_schema.md) to understand data structure
3. **Deploy the app**: See [deployment.md](../resources/deployment.md) to host online
4. **Join development**: See [development.md](../resources/development.md) to contribute features

## Need Help?

- Check the **Troubleshooting** section in the full documentation
- Review error messages in the R console
- Check that data files exist in `data/` folder
- Verify CSV format is correct (headers must match schema)

---

**Last Updated**: 2026-07-19
**Version**: 2.0 (with KVC Farm Risk Dashboard integration)
