---
title: "Quick Start Guide"
output: html_document
---

# Quick Start Guide

## Getting Started in 5 Minutes

### 1. Install Required Packages

```{r eval=FALSE}
install.packages(c("shiny", "shinydashboard", "DT", "ggplot2", "dplyr"))
```

### 2. Run the Application

```{r eval=FALSE}
# Option 1: From R console
shiny::runApp()

# Option 2: From RStudio
# Open app.R and click "Run App" button
```

### 3. Navigate the Dashboard

- **Dashboard**: View overview statistics
- **Chickens**: Manage your flock
- **Hazards**: Track and report hazards
- **About**: Learn more about the application

## Key Features at a Glance

### Managing Chickens

1. Click "Chickens" in the sidebar
2. Click "Add New Chicken" button
3. Fill in chicken details
4. Save and view in inventory table

### Reporting Hazards

1. Click "Hazards" in the sidebar
2. Click "Report New Hazard" button
3. Describe the hazard and set severity
4. Submit to add to hazard log

## Tips

- Use the refresh button to reload data
- Filter hazards by status and severity
- Check visualizations for insights
- Data is saved in CSV files in the `data/` folder

## Next Steps

For detailed information, see the complete documentation in `README.Rmd`.
